MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites \\ Supplementary Material

### 2014-12-11 chapters/MetaTracts.tex
after merge fixes
local commit

Frangi et al.~\cite{Frangi1998} introduced a process that searches for geometric structures which are tubular. They define a measure based on two geometric ratios of the second order ellipsoid given by the local Hessian matrix to measure the ``vesselness'' criterion.  In order to determine reliable Hessians, we compute 
\color{red}{ the same metric} \color{blue}{a similar metric}\color{black}. We include their work here for completeness and direct the reader to ~\cite{Frangi1998} for details. Let $\lambda_{K}$ be the eigenvalue with the $K^{th}$ smallest magnitude. Here, $|{\lambda}_{1}| \leq| {\lambda}_{2}|\leq| {\lambda}_{3}| $ are the eigenvalues of the Hessian matrix. Specifically, a pixel belonging to a vessel region will have small $\lambda_{1}$ ($|\lambda_{1}|\approx 0$) and $\lambda_{2}$, $\lambda_{3}$ of large magnitude and of equal sign($|\lambda_{1}| \ll |\lambda_{2}|$ and $|\lambda_{2}|\approx |\lambda_{3}|$). The sign indicates if the vessel is bright in a dark background or dark in a bright background. In our case the individual fibers are bright ($\lambda_2,\lambda_3 < 0$). The following measures are defined in~\cite{Frangi1998}.  

@@

\subsection{Distance based clustering}
 \label{subsec:dist_clustering}
Orientation clustering uses only orientation information. To subdivide the oriented clusters into fiber bundles, we need to include further information about the geometric proximity between MetaTracts. We use the directed Hausdorff distance for distance based clustering.
Each MetaTract is represented as a set of points ($C_P$). Formally, the directed Hausdorff distance from point set $P$ to point set $Q$ is defined as 
$H_{dir}(P,Q) = max_{p \in P} min_{q \in Q} d(p,q)$ .
The Hausdorff distance is defined as $H(P,Q) = max(H_{dir}(P,Q),H_{dir}(Q,P))$.
The Hausdorff distance is a metric so $H(P,Q) \le H(P,Q') + H(Q',Q)$ but the directed Hausdorff is not.
%
Unfortunately, the Hausdorff distance does not work well for our application since a fiber bundle may have many MetaTracts, which only partially cover the bundle (Figure~\ref{fig:length_distribution}b). If a MetaTract $P$ covers only part of the fiber bundle covered by $Q$, then $H_{dir}(P,Q)$ will be very small while $H_{dir}(Q,P)$ will be large.
%
Thus, $H(P,Q)$ will be large, even though $P$ and $Q$ are in the same fiber bundle.
Instead of using the Hausdorff distance, $\max(H_{dir}(P,Q),H_{dir}(Q,P)$, we use $\min(H_{dir}(P,Q),H_{dir}(Q,P))$. If $P$ covers only part of the fiber bundle covered by $Q$, then $\min(H_{dir}(P,Q),H_{dir}(Q,P))$ is very small.
Note that if $P$ and $Q$ overlap but do not cover the same parts of the fiber bundle, then $H_{dir}(P,Q)$ and $H_{dir}(Q,P)$ and $\min(H_{dir}(P,Q),H_{dir}(Q,P))$ will be large.
% 
The directed Hausdorff distance is very sensitive to outliers in the data.
However, because MetaTracts after orientation clustering are constructed using cylinders with similar orientations, they are not plagued by outliers.
To cluster based on MetaTract proximity, we used single linkage hierarchical clustering.
Hierarchical clustering has a single parameter $h$, the desired number of clusters.
Clusters are merged until there are only $h$ clusters left.
Hierarchical clustering is intuitive since it is easy to trace how clusters are formed and merged.
Single linkage clustering finds pairs of objects $p \in P$ and $q \in Q$ where $P \neq Q$ which are closer than other such pairs, and merges the containing clusters $P$ and $Q$.
We found that single linkage hierarchical clustering had two major drawbacks.
First, the clustering would produce some small clusters of just a few MetaTracts.
These MetaTracts were anomalies caused by overlapping fibers and did not represent true fiber bundles.
Second, if two \replacedWith{fiber bundles}{MetaTracts} were parallel for some of their length and then separated, they would sometimes be clustered into the same fiber bundle.
A short MetaTract which was parallel to both and did not extend into the separation region could form a link between the two fiber bundles, causing them to be clustered into a single bundle.
To address the problem of small clusters, we applied hierarchical clustering and then identified small clusters with few MetaTracts.
We removed the MetaTracts that were in those clusters from the data set and reapplied hierarchical clustering.
To address the problem of short MetaTracts joining different fiber bundles, we applied hierarchical clustering and then removed the
shortest tracts (tracts less than $\eta$ times the median length, set to 0.6) in each bundle. We then reapplied hierarchical clustering.
We repeated both until a steady state of clusters is reached and no new small fibers can be removed. The results of clustering the orientation clusters (see Figure~\ref{fig:orientation_clustering}a,c) are shown in Figure~\ref{fig:orientation_clustering}d,e respectively.

\subsection{Choice of clustering techniques}
\label{subsec:clus_choice}
As opposed to the two phase clustering we experimented with using proximity alone and using single step clustering of MetaTracts.
%\begin{figure}[tb]
%\centering
%		\includegraphics[width=0.4\textwidth]{images_pvis/figure7}
%		%\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
%	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
%	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
%	\caption{(A) Volume rendering of data set Fig.~\ref{fig:data-char}A.(B) shows a single slice of the data set. (C) shows two of the extracted meshes together. (D) shows the meshes separately.}
%	\label{fig:crop-16-decomp}
%\end{figure}  
%  
Figure~\ref{fig:comparison}a shows the MetaTracts that are hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters without first performing orientation clustering (ground truth is 11 clusters). As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large  incorrect clusters and small (low-cardinality) outliers. When number of clusters ($h$) is 10, two large incorrect clusters are generated and the rest are outliers. As $h$ increases, some appropriate bundles start to form. Even at $h=20$, the top fiber bundles incorrectly cluster together.
%
Figure~\ref{fig:comparison}b shows the MetaTracts, clustered using K-means clustering by proximity alone, after being embedded in a $m$-dimensional (lower) space, without a second hierarchical clustering step into 10, 15, and 20 clusters. In the lower dimensional space the spatial context is lost and fibers which are in reality far away are grouped together. Even at $h=20$, no correct bundles are identified.
\begin{figure}[tb]
	\centering
	\includegraphics[width=\linewidth]{images_pvis/comparison_all.pdf}
	\caption{Applying only (a)hierarchical and (b)\added{dimensionality reduction followed by} K-means clustering methods to MetaTracts for various numbers of clusters \added{(distance measure is minimal directed Hausdorffs.}}
	\label{fig:comparison}
	\vskip-0.2cm
\end{figure}

>>>>

Frangi et al.~\cite{Frangi1998} introduced a process that searches for geometric structures which are tubular. They define a measure based on two geometric ratios of the second order ellipsoid given by the local Hessian matrix to measure the ``vesselness'' criterion.  In order to determine reliable Hessians, we compute a similar metric. We include their work here for completeness and direct the reader to ~\cite{Frangi1998} for details. Let $\lambda_{K}$ be the eigenvalue with the $K^{th}$ smallest magnitude. Here, $|{\lambda}_{1}| \leq| {\lambda}_{2}|\leq| {\lambda}_{3}| $ are the eigenvalues of the Hessian matrix. Specifically, a pixel belonging to a vessel region will have small $\lambda_{1}$ ($|\lambda_{1}|\approx 0$) and $\lambda_{2}$, $\lambda_{3}$ of large magnitude and of equal sign($|\lambda_{1}| \ll |\lambda_{2}|$ and $|\lambda_{2}|\approx |\lambda_{3}|$). The sign indicates if the vessel is bright in a dark background or dark in a bright background. In our case the individual fibers are bright ($\lambda_2,\lambda_3 < 0$). The following measures are defined in~\cite{Frangi1998}.  

@@

\subsection{Distance based clustering}
 \label{subsec:dist_clustering}
Orientation clustering uses only orientation information. To subdivide the oriented clusters into fiber bundles, we need to include further information about the geometric proximity between MetaTracts. We use the directed Hausdorff distance for distance based clustering.
Each MetaTract is represented as a set of points ($C_P$). Formally, the directed Hausdorff distance from point set $P$ to point set $Q$ is defined as 
$H_{dir}(P,Q) = max_{p \in P} min_{q \in Q} d(p,q)$ .
The Hausdorff distance is defined as $H(P,Q) = max(H_{dir}(P,Q),H_{dir}(Q,P))$.
The Hausdorff distance is a metric so $H(P,Q) \le H(P,Q') + H(Q',Q)$ but the directed Hausdorff is not.
Unfortunately, the Hausdorff distance does not work well for our application since a fiber bundle may have many MetaTracts, which only partially cover the bundle (Figure~\ref{fig:length_distribution}b). If a MetaTract $P$ covers only part of the fiber bundle covered by $Q$, then $H_{dir}(P,Q)$ will be very small while $H_{dir}(Q,P)$ will be large.
Thus, $H(P,Q)$ will be large, even though $P$ and $Q$ are in the same fiber bundle.
Instead of using the Hausdorff distance, $\max(H_{dir}(P,Q),H_{dir}(Q,P)$, we use $\min(H_{dir}(P,Q),H_{dir}(Q,P))$. If $P$ covers only part of the fiber bundle covered by $Q$, then $\min(H_{dir}(P,Q),H_{dir}(Q,P))$ is very small.
Note that if $P$ and $Q$ overlap but do not cover the same parts of the fiber bundle, then $H_{dir}(P,Q)$ and $H_{dir}(Q,P)$ and $\min(H_{dir}(P,Q),H_{dir}(Q,P))$ will be large.
The directed Hausdorff distance is very sensitive to outliers in the data.
However, because MetaTracts after orientation clustering are constructed using cylinders with similar orientations, they are not plagued by outliers.
To cluster based on MetaTract proximity, we used single linkage hierarchical clustering.
Hierarchical clustering has a single parameter $h$, the desired number of clusters.
Clusters are merged until there are only $h$ clusters left.
Hierarchical clustering is intuitive since it is easy to trace how clusters are formed and merged.
Single linkage clustering finds pairs of objects $p \in P$ and $q \in Q$ where $P \neq Q$ which are closer than other such pairs, and merges the containing clusters $P$ and $Q$.
We found that single linkage hierarchical clustering had two major drawbacks.
First, the clustering would produce some small clusters of just a few MetaTracts.
These MetaTracts were anomalies caused by overlapping fibers and did not represent true fiber bundles.
Second, if two \replacedWith{fiber bundles}{MetaTracts} were parallel for some of their length and then separated, they would sometimes be clustered into the same fiber bundle.
A short MetaTract which was parallel to both and did not extend into the separation region could form a link between the two fiber bundles, causing them to be clustered into a single bundle.
To address the problem of small clusters, we applied hierarchical clustering and then identified small clusters with few MetaTracts.
We removed the MetaTracts that were in those clusters from the data set and reapplied hierarchical clustering.
To address the problem of short MetaTracts joining different fiber bundles, we applied hierarchical clustering and then removed the
shortest tracts (tracts less than $\eta$ times the median length, set to 0.6) in each bundle. We then reapplied hierarchical clustering.
We repeated both until a steady state of clusters is reached and no new small fibers can be removed. The results of clustering the orientation clusters (see Figure~\ref{fig:orientation_clustering}a,c) are shown in Figure~\ref{fig:orientation_clustering}d,e respectively. 
\subsection{Choice of clustering techniques}
\label{subsec:clus_choice}
As opposed to the two phase clustering we experimented with using proximity alone and using single step clustering of MetaTracts.
%\begin{figure}[tb]
%		\includegraphics[width=0.4\textwidth]{images_pvis/figure7}
%		%\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
%	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
%	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
%	\caption{(A) Volume rendering of data set Fig.~\ref{fig:data-char}A.(B) shows a single slice of the data set. (C) shows two of the extracted meshes together. (D) shows the meshes separately.}
%	\label{fig:crop-16-decomp}
Figure~\ref{fig:comparison}a shows the MetaTracts that are hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters without first performing orientation clustering (ground truth is 11 clusters). As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large  incorrect clusters and small (low-cardinality) outliers. When number of clusters ($h$) is 10, two large incorrect clusters are generated and the rest are outliers. As $h$ increases, some appropriate bundles start to form. Even at $h=20$, the top fiber bundles incorrectly cluster together.
Figure~\ref{fig:comparison}b shows the MetaTracts, clustered using K-means clustering by proximity alone, after being embedded in a $m$-dimensional (lower) space, without a second hierarchical clustering step into 10, 15, and 20 clusters. In the lower dimensional space the spatial context is lost and fibers which are in reality far away are grouped together. Even at $h=20$, no correct bundles are identified.
\begin{figure}[tb]
	\centering
	\includegraphics[width=\linewidth]{images_pvis/comparison_all.pdf}
	\caption{Applying only (a)hierarchical and (b)\added{dimensionality reduction followed by} K-means clustering methods to MetaTracts for various numbers of clusters \added{(distance measure is minimal directed Hausdorffs.}}
	\label{fig:comparison}
	\vskip-0.2cm
\end{figure}
